\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-03-17 doc/paper.Rmd
Chumbley score sketch

The algorithm works in two phases, namely, an optimization step and a validation step.

In the optimization step, the two markings are aligned horizontally such that within a pre-defined window of length $w_o$ the correlation between $x(t_1)$ and $y(t_2)$ is maximized:
\[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
\]
This results in an optimal vertical (in-phase) shift of $t_1^o - t_2^o$ for aligning the two markings.

@@

```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}

@@

A pre-processing step to the algorithm is to choose a coarseness value which  is used as a parameter to the lowess smoothing function. The coarseness essentially gives the proportion of points which influence the smooth at each value, which means larger values lead to more smoothness. The lowess smoothing is applied to each of two sets of vectorized striae or marks $x(t_1)$ and $y(t_2)$, before proceeding to the algorithm.

>>>>

In a pre-processing step  the two markings are smoothed using a lowess \citep{lowess} with coarseness parameter $c$. The purpose of this smoothing is to remove drift and (sub)class characteristics.

The Chumbley scores is calculated in two phases, namely, an optimization step and a validation step.
In the optimization step, the two markings are aligned horizontally such that within a pre-defined window of length $w_o$ the correlation between $x(t_1)$ and $y(t_2)$ is maximized:
\[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
\]
This results in an optimal vertical (in-phase) shift of $t_1^o - t_2^o$ for aligning the two markings.

@@

```{r win-comparison, warning=FALSE, fig.width=8, fig.height=3.5, fig.cap="In- and out-of-phase samples from two markings created by the same source. On the left, sample windows are in-phase, on the right the pairs of sampled windows are out-of-phase. The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets."}
sigs_graphics <- readRDS("../data/sigs_generate_window.rds")

# Matching signature
d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==276,arr.ind = TRUE),]
data1<-d1 %>%select(l30)
#data1<- data1$y
data2<- d2 %>%select(l30)
#data2<- data2$y
window_opt = 120 
window_val = 50
coarse = 1
data1<- matrix(unlist(data1))
data2<- matrix(unlist(data2))

sigs1.min = sigs_graphics[[1]] %>% filter(land_id==1) %>% select(y) %>% min()
sigs2.min = sigs_graphics[[1]] %>% filter(land_id==276) %>% select(y) %>% min()
window_val <- 2*window_val*0.645

same.shift <- -(365-338)*1/.645 + sigs2.min - sigs1.min 
xs <- seq(450, 2000, length=10)
rects1 <- data.frame(ymin=-2.5, ymax= Inf, xmin = xs, xmax = xs+window_val)
rects2 <- data.frame(ymin=-Inf, ymax=-3.5, xmin = xs+same.shift, xmax = same.shift+xs+window_val)
rectsSS <- data.frame(rbind(rects1, rects2), type=rep(1:2, each=length(xs)))

p1 <- sigs_graphics[[1]] %>% filter(land_id %in% c(1, 276)) %>% 
  ggplot() + 
  geom_segment(aes(x = xmin, y=ymin, xend = xmin, yend=ymax), 
               colour="grey50", linetype=3, data = rectsSS) +
  geom_segment(aes(x = xmax, y=ymin, xend = xmax, yend=ymax), 
               colour="grey50", linetype=3, data = rectsSS) +
  geom_segment(aes(x = xmin, y=ymin, xend = xmin+same.shift, yend=ymin-1), 
               colour="grey50", linetype=3, data = rectsSS %>% filter(type==1)) +
  geom_segment(aes(x = xmax, y=ymin, xend = xmax+same.shift, yend=ymin-1), 
               colour="grey50", linetype=3, data = rectsSS %>% filter(type==1)) +
  geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, group=type), 
            data=rectsSS, fill= "grey90") +
  geom_line(aes(x = y, y=l30-7*I(land_id==276), group=land_id)) + 
  theme_bw() +
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  xlab("Relative Location (in microns)") +
  ggtitle("In-phase sample")

diff.shift <- -(365-338)*1/.645 + sigs2.min - sigs1.min + 50 
rects2 <- data.frame(ymin=-Inf, ymax=-3.5, xmin = xs+diff.shift, xmax = diff.shift+xs+window_val)
rectsDS <- data.frame(rbind(rects1, rects2), type=rep(1:2, each=length(xs)))

p2 <- sigs_graphics[[1]] %>% filter(land_id %in% c(1, 276)) %>% 
  ggplot() + 
  geom_segment(aes(x = xmin, y=ymin, xend = xmin, yend=ymax), 
               colour="grey50", linetype=3, data = rectsDS) +
  geom_segment(aes(x = xmax, y=ymin, xend = xmax, yend=ymax), 
               colour="grey50", linetype=3, data = rectsDS) +
  geom_segment(aes(x = xmin, y=ymin, xend = xmin+diff.shift, yend=ymin-1), 
               colour="grey50", linetype=3, data = rectsDS %>% filter(type==1)) +
  geom_segment(aes(x = xmax, y=ymin, xend = xmax+diff.shift, yend=ymin-1), 
               colour="grey50", linetype=3, data = rectsDS %>% filter(type==1)) +
  geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax, group=type), 
            data=rectsDS, fill= "grey90") +
  geom_line(aes(x = y, y=l30-7*I(land_id==276), group=land_id)) + 
  theme_bw() +
  theme(axis.title.y = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  xlab("Relative Location (in microns)") +
  ggtitle("Out-of-phase sample")

library(gridExtra)
grid.arrange(p1,p2, ncol=2)
```

```{r win-comparison-2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.", eval=FALSE}

@@

### 2018-03-17 doc/paper.tex
Chumbley score sketch

The algorithm works in two phases, namely, an optimization step and a
validation step.

In the optimization step, the two markings are aligned horizontally such
that within a pre-defined window of length \(w_o\) the correlation
between \(x(t_1)\) and \(y(t_2)\) is maximized: \[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
\] This results in an optimal vertical (in-phase) shift of
\(t_1^o - t_2^o\) for aligning the two markings.

@@

\caption{ The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.}\label{fig:win-comparison}
\end{figure}

@@

A pre-processing step to the algorithm is to choose a coarseness value
which is used as a parameter to the lowess smoothing function. The
coarseness essentially gives the proportion of points which influence
the smooth at each value, which means larger values lead to more
smoothness. The lowess smoothing is applied to each of two sets of
vectorized striae or marks \(x(t_1)\) and \(y(t_2)\), before proceeding
to the algorithm.

>>>>

In a pre-processing step the two markings are smoothed using a lowess
\citep{lowess} with coarseness parameter \(c\). The purpose of this
smoothing is to remove drift and (sub)class characteristics.

The Chumbley scores is calculated in two phases, namely, an optimization
step and a validation step. In the optimization step, the two markings
are aligned horizontally such that within a pre-defined window of length
\(w_o\) the correlation between \(x(t_1)\) and \(y(t_2)\) is maximized:
\[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x (t_1, t_1 + w_o), y(t_2, t_2 + w_o) \right)
\] This results in an optimal vertical (in-phase) shift of
\(t_1^o - t_2^o\) for aligning the two markings.

@@

\caption{In- and out-of-phase samples from two markings created by the same source. On the left, sample windows are in-phase, on the right the pairs of sampled windows are out-of-phase. The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets.}\label{fig:win-comparison}
\end{figure}

@@
